%% Modified from original file `template.tex'.
%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.


\documentclass[11pt,letterpaper,roman]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')

% moderncv themes
\moderncvstyle{classic}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvcolor{grey}                              % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
%\renewcommand{\familydefault}{\sfdefault}         % to set the default font; use '\sfdefault' for the default sans serif font, '\rmdefault' for the default roman one, or any tex font name
%\nopagenumbers{}                                  % uncomment to suppress automatic page numbering for CVs longer than one page

% character encoding
\usepackage[utf8]{inputenc}

% adjust the page margins
\usepackage[letterpaper,scale={0.75,0.8}]{geometry}

% font
\usepackage{tgpagella}
\renewcommand{\familydefault}{\sfdefault}         

\renewcommand*{\addresssymbol}       {}
\renewcommand*{\mobilephonesymbol}   {}
\renewcommand*{\emailsymbol}         {}
\renewcommand*{\homepagesymbol}      {}
\renewcommand*{\addressfont}{\small\mdseries}

% personal data
\name{Michael C.}{Hughes}
\phone[mobile]{+1~(617)~371~7611}
\email{mike@michaelchughes.com}
\homepage{www.michaelchughes.com}
%\extrainfo{additional information}                 % optional, remove / comment the line if not wanted
%\photo[64pt][0.4pt]{picture}                       % optional, remove / comment the line if not wanted; '64pt' is the height the picture must be resized to, 0.4pt is the thickness of the frame around it (put it to 0pt for no frame) and 'picture' is the name of the picture file

%----------------------------------------------------------------------------------
%            content
%----------------------------------------------------------------------------------
\begin{document}
%-----       letter       ---------------------------------------------------------
% recipient data
\recipient{Harvard University CS Faculty Search Committee}{
Paulson School of Engineering and Applied Sciences\\
29 Oxford Street\\
Cambridge, MA 02138}
\date{\today}
\opening{Dear Members of the Faculty Search Committee,}
\closing{Sincerely,}

\makelettertitle

I am writing to apply for a tenure-track assistant professorship at Harvard SEAS. I believe my background in statistical machine learning and its applications to healthcare make me a good fit for the advertised position within the Computer Science area.

I am currently a postdoctoral fellow at Harvard SEAS, after completing my Ph.D. in Computer Science from Brown University.
My research develops machine learning methods
that discover actionable knowledge from large, messy datasets. 
My focus is healthcare applications, where new probabilistic models and optimization algorithms are needed to learn from abundant yet noisy and unlabeled electronic health records to produce insights clinicians can trust.
In one project, we modeled common trajectories in the Intensive Care Unit (ICU) to predict hourly need for interventions like mechanical ventilation. In ongoing work with Massachusetts General Hospital (MGH), we are discovering subpopulations of major depression disorder that require distinct drug combination therapies. Our efforts build on the core contributions of my Ph.D. thesis: reliable algorithms for clustering hierarchical or sequential datasets which can scale up to cover millions of news articles or the whole human genome.
I strive to deliver results beyond top-tier publications, such as my open-source BNPy toolbox (used by data scientists at the NY Times). As a future faculty member, I hope to build a group known for strong methodological contributions within machine learning as well as successful deployment of clinical decision support systems integrated at the point-of-care.

I am excited to join Harvard as a faculty member to pursue this vision.
Within CS, I would continue my close collaboration with Prof. Doshi-Velez and also seek projects with other experts in artificial intelligence, human-computer interaction, and systems to tame large-scale multimodal health datasets. 
I see opportunities to work with experts in the Medical School and School of Public Health to bring the benefits of machine learning to patients and co-teach interdisciplinary courses. I could partner with experts in Statistics or Psychology to explore topics in explainable AI such as interpretability and causality.
I would gladly engage with efforts in CRCS, IACS, and the Data Science Initiative to promote data science across departments.
 Finally, 
I have significant experience in K-12 STEM outreach (2 years as a Microsoft TEALS ``CS in every High School'' instructor), and I plan to continue efforts to improve STEM diversity from K-12 to graduate school.
I am eager to join Harvard's faculty, and I look forward to hearing from you.

\makeletterclosing

\end{document}


%% end of file `template.tex'.
